Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-02-15 bbbq_article.tex
Rewrite of abstract, progress #179

The human immune system is unexpectedly attentive towards
human transmembrane helices (TMHs), as TMH-derived peptide fragments
bind to histocompatibility complex (MHC) class I more often then expected
by chance only, which can be mostly explained by the low hydrophobicity
of those residues.
The physiological reason is yet unclear,
yet this finding hints that there has been selection upon detecting TMH-derived
peptides in pathogens.
If this signature of selection is present in MHC-II is unknown.
This study shows that MHC-I also has more
epitopes derived from a TMH for a pathogen proteome, when compared with
a host proteome.
Additionally, MHC-II binds to peptides derived from TMHs 
more often than expected by chance.
Lastly, we show the TMHs are evolutionarily more conserved,
as less SNPs are found in TMH than expected by chance.
Our findings suggest that the immune system is more
vigilant to TMHs than expected by chance
and suggests that an evolutionary signal in the evolutionary history of a pathogen
may be found.

>>>>

Cytolytic T cell responses are predicted to be biased towards membrane proteins. 
Because the peptide-binding grooves of most haplotypes 
of histocompatibility complex class I (MHC-I) are relatively hydrophobic, 
peptide fragments derived from transmembrane helices (TMHs) 
are predicted to be presented more often than expected based on their abundance.
However, the physiological reason of why membrane proteins might be 
over-presented is unclear
In this study, we show the TMHs are evolutionarily more conserved, 
as relatively less single nucleotide polymorphisms (SNPs) 
are present in TMH-coding chromosomal regions 
compared to regions coding for extracellular and cytosolic protein regions. 
Moreover, we show that the over-presentation of TMH-derived peptides is 
likely general, as it is predicted for diverse microbial pathogens 
and for both MHC-I and MHC-II.
Thus, our findings suggest that T cells might respond 
more to membrane proteins, because these are evolutionary more conserved.
We speculate that TMHs might be less prone of the escape mutations 
that enable pathogens to evade T cell responses.
